---
title: "Italian provinces (own calculations) - work in progress (DRAFT)"
author: "Marcelo Eduardo Redoschi"
output: html_document
---

## 1. Initial setup 

### Libraries and paths 

```{r}
library(readr)
library(lubridate)
library(dplyr)
library(markdown)
source("helper_functions.R")
datasets_path<-file.path(getwd(),"..", "COVID-19") 
```

### Date functions 

```{r}
today<-today() 
yesterday<-yesterday()
two_days_ago<-two_days_ago() 
three_days_ago<-three_days_ago()
```

## 2. Provincial data - CSV Import step

*dpc-covid19-ita-province.csv*
```{r}
provinces_csv_fname<-file.path(datasets_path,"dati-province","dpc-covid19-ita-province.csv")

# na.string explicitly set to something else, since NA in this file means the Naples province abbreviation!
provinces_raw<-read.csv(provinces_csv_fname,encoding="UTF-8", na.strings="undefined") 


```

## 3. Filter out rows marked "In fase di definizione/aggiornamento"

```{r}
to_be_determined_or_updated<-"In fase di definizione/aggiornamento"
provinces_it<-provinces_raw[provinces_raw$denominazione_provincia!=to_be_determined_or_updated,]
```

## 4. Italian administrative regions
```{r}
italian_adm_regions<-distinct(select(provinces_it, denominazione_regione, codice_regione))
col_names<-colnames(italian_adm_regions)
intl_col_names<-gsub("denominazione_regione","region_name",col_names)
intl_col_names2<-gsub("codice_regione","region_code",intl_col_names)
colnames(italian_adm_regions)<-intl_col_names2 
rm(intl_col_names)
rm(intl_col_names2)
italian_adm_regions<-arrange(italian_adm_regions, region_name)
```

## 5. Italian provinces (belong to one region and contain more than one city or municipality)
```{r}
italian_adm_provinces<-distinct(select(provinces_it, denominazione_provincia, codice_provincia, sigla_provincia, denominazione_regione, codice_regione, lat, long))
col_names<-colnames(italian_adm_provinces)
intl_col_names<-gsub("denominazione_provincia","province_name",col_names)
intl_col_names2<-gsub("codice_provincia","province_code",intl_col_names)
intl_col_names3<-gsub("sigla_provincia","province_abbreviation",intl_col_names2)
intl_col_names4<-gsub("denominazione_regione","region_name",intl_col_names3)
intl_col_names5<-gsub("codice_regione","region_code",intl_col_names4)

colnames(italian_adm_provinces)<-intl_col_names5 
rm(intl_col_names)
rm(intl_col_names2)
rm(intl_col_names3)
rm(intl_col_names4)
rm(intl_col_names5)
italian_adm_provinces<-arrange(italian_adm_provinces,  province_name)
# should have zero observations, if Naples was imported correctly.
# provinces_without_abbreviation<-italian_adm_provinces[is.na(italian_adm_provinces$province_abbreviation),] 
```

## 6. Province case data


### Initial preparation 

```{r}
provinces<-select(provinces_it, data, codice_provincia, totale_casi)
provinces$dt<-as.Date(ymd_hms(provinces$data))
provinces$data<-NULL 
col_names<-colnames(provinces)
intl_col_names<-gsub("codice_provincia","province_code",col_names)
intl_col_names2<-gsub("totale_casi","total_cases",intl_col_names)
colnames(provinces)<-intl_col_names2
```

### Today 

```{r}
prov_today<-provinces[provinces$dt==today,]
prov_today$dt<-NULL
prov_today$current_total<-prov_today$total_cases 
prov_today$total_cases<-NULL 
```

### Yesterday 

```{r}
prov_yesterday<-provinces[provinces$dt==yesterday,]
prov_yesterday$dt<-NULL
prov_yesterday$yesterday_total_cases<-prov_yesterday$total_cases 
prov_yesterday$total_cases<-NULL
```

### Join
```{r}
prov<-inner_join(prov_yesterday, prov_today)
prov$todays_rate<-prov$current_total/prov_yesterday$yesterday_total_cases
```

### Two days ago

```{r}
prov_2_days_ago<-provinces[provinces$dt==two_days_ago,]
prov_2_days_ago$dt<-NULL
prov_2_days_ago$total_two_days_ago<-prov_2_days_ago$cases
col_names<-colnames(prov_2_days_ago)
revised_col_names<-gsub("total_cases","total_cases_two_days_ago",col_names)
colnames(prov_2_days_ago)<-revised_col_names
```

### Three days ago

```{r}
prov_3_days_ago<-provinces[provinces$dt==three_days_ago,]
prov_3_days_ago$dt<-NULL
prov_3_days_ago$total_two_days_ago<-prov_3_days_ago$cases
col_names<-colnames(prov_3_days_ago)
revised_col_names<-gsub("total_cases","total_cases_three_days_ago",col_names)
colnames(prov_3_days_ago)<-revised_col_names
```

```{r}
provinces_calc<-inner_join(prov_today, prov_yesterday)
provinces_calc<-inner_join(provinces_calc, prov_2_days_ago)
provinces_calc<-inner_join(provinces_calc, prov_3_days_ago)
provinces_calc$current_rate<-provinces_calc$current_total/provinces_calc$yesterday_total_cases
provinces_calc$yesterdays_rate<-provinces_calc$yesterday_total_cases/provinces_calc$total_cases_two_days_ago
provinces_calc$rate_two_days_ago<-provinces_calc$total_cases_two_days_ago/provinces_calc$total_cases_three_days_ago
province_names<-select(italian_adm_provinces, province_code, province_name, province_abbreviation)
provinces_calc<-inner_join(province_names, provinces_calc)
rm(province_names)
```

```{r}
write.csv(provinces_calc,"provinces_calc.csv")
```
*Last updated: 31/3/2020*